---
title: "R Notebook"
output: html_notebook
---

# Load in Libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Load in DF
```{r}
public_housing <- read.csv("C:\\Users\\ethan\\OneDrive\\Documents\\STOR320\\finalDF.csv")
```

# Make a new DF for Scatter Plotting
```{r}
scatter_data <- public_housing %>%
  group_by(State) %>%
  summarise_all(mean, na.rm = TRUE) %>%
  select(STATE2KX, SPENDING_PER_MONTH, Republican_Percent, Moderate_Percent, Democrat_Percent, State) %>% 
  mutate(State = tolower(State)) %>%
  mutate(Dominant_Party = case_when(
    Republican_Percent > Moderate_Percent & Republican_Percent > Democrat_Percent ~ "Republican",
    Moderate_Percent > Republican_Percent & Moderate_Percent > Democrat_Percent ~ "Moderate",
    Democrat_Percent > Republican_Percent & Democrat_Percent > Moderate_Percent ~ "Democrat",
    TRUE ~ "Equal"  # If all percentages are equal
  )) %>%
  filter(STATE2KX <= 51)

  #gather(key = "Party", value = "value", Republican_Percent, Moderate_Percent, Democrat_Percent) %>%

head(scatter_data, 60)
#nrow(scatter_data)
```

# Making Scatter plot (TAKING INTO ACCOUNT POPULATION SIZE DIFFRENCES THIS IS LOWKEY CRAZY)
```{r}
ggplot(scatter_data, aes(x = STATE2KX, y = SPENDING_PER_MONTH, size = SPENDING_PER_MONTH, color = Dominant_Party)) +
  geom_point() +
  labs(x = "State", y = "Government Spending", color = "Dominant Party") +
  scale_color_manual(values = c("Republican" = "red", "Moderate" = "gray", "Democrat" = "blue", "Equal" = "purple")) +
  ggtitle("Government Spending vs. State with Dominant Party Affiliation")
```

```{r}
scatter_data2 <- scatter_data %>%
  filter(Dominant_Party != "Equal")

ggplot(scatter_data2, aes(x = STATE2KX, y = SPENDING_PER_MONTH, size = SPENDING_PER_MONTH, color = Dominant_Party)) +
  geom_point() +
  geom_smooth(aes(group = Dominant_Party), method = "lm", se = FALSE) +
  labs(x = "State", y = "Government Spending", color = "Dominant Party") +
  scale_color_manual(values = c("Republican" = "red", "Moderate" = "gray", "Democrat" = "blue", "Equal" = "purple")) +
  ggtitle("Government Spending vs. State with Dominant Party Affiliation")

```


# Aditi Idea: First Merge Scatter data and mappable data
## Map to show trend (State Color is political lean) - dot over is gov't spending
```{r}
states <- map_data("state")
mappable_data <- merge(scatter_data, states, by.x = "State", by.y = "region", all = TRUE, na.rm=FALSE) %>%
  select(-subregion, -STATE2KX)

head(mappable_data)
```
# Map it out
```{r}
ggplot(mappable_data, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = Dominant_Party), color = "black") + 
  coord_map() +
  scale_fill_manual(values = c("Republican" = "red", "Moderate" = "gray", "Democrat" = "blue", "Equal" = "purple")) +
  ggtitle("Party Affiliation by State")
```

```{r}
ggplot(mappable_data, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = SPENDING_PER_MONTH), color = "black") + 
  coord_map() +
  scale_fill_gradient(low = "white", high = "red") +
  ggtitle("Spending by State")
```


# Lowkey this graph just didnt turn out
```{r}
ggplot(mappable_data, aes(long, lat)) +
  geom_polygon(aes(group = group, fill = SPENDING_PER_MONTH, color = Dominant_Party), size = 1) +
  scale_color_manual(values = c("Republican" = "red", "Moderate" = "gray", "Democrat" = "blue", "Equal" = "purple")) +
  scale_fill_gradient(low = "white", high = "black") +
  coord_map() +
  theme_minimal() +
  labs(title = "Spending per State")
```

# With more time maybe do the dot but thats gonna take me making a new df and merging it back in I think