---
title: "Homework 1"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Instructions

**Exercises:** 1-5 (Pgs. 6-7); 1-2, 5 (Pg. 12); 1-5 (Pgs. 20-21); Open Response

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
```{r}
ggplot(data=mpg)
```

I see absolutely nothing. There is just a blank space for a graph. Why am I even doing this nonsense?

## Exercise 2
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```


The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```

Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
```{r}
ggplot(data = mpg) +  geom_point(    mapping = aes(x = displ, y = hwy), color = "blue"  )
```

The closed parenthesis ")" was missplaced.

##  Exercise 2
```{r}
#?mpg
```
Catagorical Variables: manufacturer, model, year, trans, drv, fl, class

##  Exercise 5
```{r}
#?geom_point
```
Changes the stroke weight of each border.
# Chapter 1 (Pgs. 20-21)

##  Exercise 1
```{r}
#
```

What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
Line: geom_smooth
box: geom_boxplot
histogram: geom_histogram
area chart: geom_area

##  Exercise 2
```{r}
ggplot(  data = mpg,  mapping = aes(x = displ, y = hwy, color = drv) ) +  geom_point() +  geom_smooth(se = FALSE)
```

##  Exercise 3
```{r}
#
```

It removes the legend. Maybe because we already knew what you were looking at on those graphs as we've already looked at that info the graph above.

##  Exercise 4
```{r}
#
```
Makes a line over the data to help the eye find trends.
##  Exercise 5

I'm pretty sure they will look the same

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.

```{r}
#?diamonds
ggplot(data = diamonds, mapping = aes(x = color, y = price)) + 
  geom_boxplot() + facet_wrap(vars(cut))
```


## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 
```{r}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + geom_point(mapping = aes(color = color, shape = cut))
```

